High-resolution ALMA observations of V4046\,Sgr: a circumbinary disc with a thin ring

### 2021-05-31 Draft_V4046Sgr_mnras.tex
Updates from Overleaf

The chosen regularization term $S$ for this case was the standard image entropy, or
\begin{equation}
S = \sum_{j=1}^M \frac{I_j^m}{M} \ln\left(\frac{I_j^m}{M}\right),
\end{equation}
where $M$ is the default pixel intensity value and is set to $10^{-3}$ times the theoretical noise of the dirty map (as inferred from the visibility weights $\omega_k$). Here we set $\lambda = 0.01$. Similar applications of {\tt uvmem} in the context of protoplanetary discs can be found, for example, in \citet{Casassus2013Natur, 2018MNRAS.477.5104C, Casassus2019MNRAS.483.3278C, Perez2019AJ....158...15P} and in \citet{2020ApJ...889L..24P} using long baseline data. An advantage of {\tt uvmem} compared to more traditional imaging strategies, such as provided by the {\tt tclean} task in CASA, is that the effective angular resolution of the model image is $\sim$3 times finer than the natural weights clean beam \citep[][]{2018A&C....22...16C}, giving us a new approximate {\tt uvmem} resolution of $\sim$0$\farcs021 \,\times \,\sim$0$\farcs018$. This angular resolution is comparable to uniform or super-uniform weights in {\tt tclean}, but it preserves the natural-weights point-source sensitivity.

The resulting {\tt uvmem} image shown in the top right panel of Fig.~\ref{fig:images_vs_simulated} exhibits previously unseen substructure of the disc. This image reveals that the disc features two rings of large dust grains with a broad gap between them, i.e. Ring13 at 13\,au and Ring24 starting at 24\,au. The wide and bright Ring24 reaches its peak intensity at $\sim$30\,au, continues relatively flat and then breaks at $\sim$35\,au into a steeper tail. While this is the first observation of Ring13, \citet{Ru_z_Rodr_guez_2019} did anticipate its existence as their ALMA continuum image showed a distinct excess between $\sim$10 and 17\,au. The {\tt uvmem} map has a resolution of 21$\times$18\,mas and an RMS noise of 26$\micron$Jy per resolution beam, at a frequency of 237 GHz.

@@

The ring width can be measured in polar coordinates by fitting 1-D Gaussians at each azimuth, thus obtaining a width and centroid. The top panel of Fig.~\ref{fig:polarring} shows the positions of the obtained centroids and widths over the polar decomposition of the ALMA image. On average, we obtained a FWHM of 2.83$\pm$0.50\,au, and a stellocentric radius of 13.15$\pm$0.42\,au. As the {\tt uvmem} model image has an approximate {\tt uvmem} beam of $\sim$0$\farcs021 \,\times \,\sim$0$\farcs018$ (or $\sim$1.4\,au at 71.1\,pc), we see that Ring13 is resolved. After subtraction of the {\tt uvmem} beam, the ring width is $\sim$2.46$\pm$0.56\,au.

@@

>>>>

The chosen regularization term $S$ for this case was the standard image entropy, or
\begin{equation}
S = \sum_{j=1}^M \frac{I_j^m}{M} \ln\left(\frac{I_j^m}{M}\right),
\end{equation}
where $M$ is the default pixel intensity value and is set to $10^{-3}$ times the theoretical noise of the dirty map (as inferred from the visibility weights $\omega_k$). Here we set $\lambda = 0.01$. Similar applications of {\tt uvmem} in the context of protoplanetary discs can be found, for example, in \citet{Casassus2013Natur, 2018MNRAS.477.5104C, Casassus2019MNRAS.483.3278C, Perez2019AJ....158...15P} and in \citet{2020ApJ...889L..24P} using long baseline data. An advantage of {\tt uvmem} compared to more traditional imaging strategies, such as provided by the {\tt tclean} task in CASA, is that the effective angular resolution of the model image is $\sim$3 times finer than the natural weights clean beam \citep[][]{2018A&C....22...16C}, giving us a new approximate {\tt uvmem} resolution of $\sim$0$\farcs021 \,\times \,\sim$0$\farcs018$. This angular resolution is comparable to uniform or super-uniform weights in {\tt tclean}, but it preserves the natural-weights point-source sensitivity. The RMS noise of the {\tt uvmem} image is 26\,$\micron$Jy per resolution beam, at a frequency of 237\,GHz.

The resulting {\tt uvmem} image shown in the top right panel of Fig.~\ref{fig:images_vs_simulated} exhibits previously unseen substructure of the disc. This image reveals that the disc features two rings of large dust grains with a broad gap between them, i.e. Ring13 at 13\,au and Ring24 starting at 24\,au. The wide and bright Ring24 reaches its peak intensity at $\sim$30\,au, continues relatively flat and then breaks at $\sim$35\,au into a steeper tail. While this is the first observation of Ring13, \citet{Ru_z_Rodr_guez_2019} did anticipate its existence as their ALMA continuum image showed a distinct excess between $\sim$10 and 17\,au.

@@

@@

The width of Ring13 can be measured in polar coordinates by fitting 1-D Gaussians at each azimuth, thus obtaining a width and centroid. The top panel of Fig.~\ref{fig:polarring} shows the positions of the obtained centroids and widths over the polar decomposition of the ALMA image. On average, we obtained a FWHM of 2.83$\pm$0.50\,au, and a stellocentric radius of 13.15$\pm$0.42\,au. As the {\tt uvmem} model image has an approximate {\tt uvmem} beam of $\sim$0$\farcs021 \,\times \,\sim$0$\farcs018$ (or $\sim$1.4\,au at 71.1\,pc), we see that Ring13 is resolved. After subtraction of the {\tt uvmem} beam, the ring width is $\sim$2.46$\pm$0.56\,au.
